%
% Copyright © 2018 Peeter Joot.  All Rights Reserved.
% Licenced as described in the file LICENSE under the root directory of this GIT repository.
%
%{
   \chapter{Electromagnetism.}
      \section{Conventional formulation.}
         \input{freespace.tex}
         \subsection{Problems.}
            \input{landauEasy.tex}
      \section{Maxwell's equation.}
         \input{isotropicMaxwells.tex}
      \section{Wave equation.}
         \input{waveequation.tex}
      %\section{Continuity equations.}
         \input{continuity.tex}
      \section{Dynamics.}
         \subsection{Inverting Maxwell's equation.}
            \input{jefimenkosEquations.tex}
      \section{Statics.}
         \subsection{Inverting the Maxwell statics equation.}
            \input{statics.tex}
         \subsection{Example.  Infinite line charge and current.}
            \input{statics_infiniteLineCharge.tex}
         \subsection{Example.  Infinite planar charge and current.}
            \input{statics_infinitePlane.tex}
         \subsection{Example.  Field of a ring of charge or current density.}
            \input{ringField.tex}
      \section{Energy and Momentum.}
         \subsection{Field energy and momentum density and the energy-momentum tensor.}
            \input{poyntingF.tex}
         \subsection{Poynting's theorem.}
            \input{poyntingTheorem.tex}
         \subsection{Relation to Lorentz force.}
            \input{poyntingLorentzForce.tex}
         \subsection{Examples: Some static fields.}
            \input{poyntingStaticExamples.tex}
         \subsection{Complex energy and power.}
            \input{poyntingFComplexPower.tex}
      \section{Plane waves.}
         \input{planewaves.tex}
      \section{Polarization.}
         \subsection{Plane wave.}
            \input{polarization.tex}
         \subsection{Circular polarization basis.}
            \input{polarization_circular.tex}
         \subsection{Linear polarization.}
            \input{polarization_linearPolarization.tex}
         \subsection{Other phase dependence and energy momentum.}
            \input{polarization_phaseAndEnergyMomentum.tex}
         \subsection{Elliptical parameterization.}
            \input{polarization_elliptical.tex}
         \subsection{Pseudoscalar imaginary.}
            \input{polarization_pseudoscalarImaginary.tex}
         \subsection{Problems.}
            \input{polarization_problems.tex}
      \section{Transverse fields in a waveguide.}
         \input{transverseField.tex}
      \section{Boundary value conditions.}
         \input{boundarySurfaceSources.tex}
      \section{Multivector potential.}
         \subsection{General potential representation.}
            \input{potentialSection.tex}
         \subsection{Electric sources.}
            \input{potentialSection_electric.tex}
         \subsection{Magnetic sources.}
            \input{potentialSection_magnetic.tex}
         \subsection{Far field.}
            \input{potentialSection_farfield.tex}
         \subsection{Gauge transformations}
            \input{potentialSection_gauge.tex}
         \subsection{Lorenz gauge}
            \input{potentialSection_LorenzGauge.tex}
      \section{Lorentz force.}
         \subsection{GA statement.}
            \input{lorentzForce.tex}
         \subsection{Constant magnetic field.}
            \input{lorentzForce_constantMagnetic.tex}
      \chapter{Rework.}
      \section{Electrostatics.}
         \input{electrostatics.tex}
         \subsection{Enclosed charge.}
            \input{electrostatics_enclosedCharge.tex}
         \subsection{Electric potential.}
            \input{electrostatics_electricPotential.tex}
         \subsection{Inverting the gradient equations.}
            \input{electrostatics_invertingGradient.tex}
         \subsection{Poisson equation solution.}
            \input{electrostatics_poissonEquationSolution.tex}
         \subsection{Example: Straight line charge.}
            \input{linecharge.tex}
         \subsection{Example: Circular line charge.}
            \input{circularlinecharge.tex}
         \subsection{Problems.}
            \input{electrostaticsProblems.tex}
      \section{Magnetostatics.}
         \input{magnetostatics.tex}
         \subsection{Vector potential.}
            \input{magnetostatics_vectorPotential.tex}
         \subsection{Enclosed current density.}
            \input{magnetostatics_enclosedCurrentDensity.tex}
         \subsection{Enclosed current.}
            \input{magnetostatics_enclosedCurrent.tex}
         \subsection{Biot-Savart law.}
            \input{magnetostatics_biotSavart.tex}
         \subsection{Example.  Ampere's law for magnetic field between two current sources.}
            \input{amperesLawMagnetostaticsExample.tex}
      \section{Dielectric and magnetic media.}
         \input{inMatter.tex}
%   \section{Radiation and scattering}
%TODO.
   %   \subsection{Problem solutions}
   %      \shipoutAnswer

%\part{First draft fragments to rewrite.}
%   \chapter{Maxwell's equations}
%      \input{maxwellsEquations.tex}
%      \section{Potentials}
%         \input{potentialSection.tex}
%      \section{Problem solutions}
%         \shipoutAnswer
%%   \chapter{Time harmonic fields}
%%      \section{Frequency domain}
%%         \input{../frequencydomain/frequencydomainMaxwells.tex}
%%      \section{Plane waves}
%%         \input{frequencydomainPlaneWaves.tex}
%%      \section{Problem solutions}
%%         \shipoutAnswer
%   \chapter{Polarization}
%      \input{working/ellipticalWavesPhasor.tex}
%      \section{Problem solutions}
%         \shipoutAnswer
%%</TEMPORARY COMMENT OUT FOR QUICKER BUILD.>
%}
