%
% Copyright © 2018 Peeter Joot.  All Rights Reserved.
% Licenced as described in the file LICENSE under the root directory of this GIT repository.
%
%{
\chapter{Geometric Algebra.}
   \section{Prerequisites.}
      \input{prerequisites.tex}
      \subsection{Vector space.}
         \input{prereq_vectorspace.tex}
         % problems
         \input{prereq_vectorspace_Rn.tex}
         \input{functionvectorspace.tex}
         \input{paulivectorspace.tex}
      \subsection{Basis, span and dimension.}
         \input{prereq_basisspandim.tex}
      \subsection{Standard basis, length and normality.}
         \input{prereq_standardbasis.tex}
   \section{Multivectors.}
      \input{multivector.tex}
   \section{Nomenclature.}
      \input{multivector_nomenclature.tex}
         %\input{analysis_cavaet.tex}
   \section{Colinear vectors.}
      \input{colinearVectors.tex}
   \section{Normal vectors.}
      \input{normalVectors.tex}
   \section{2D multiplication table.}
      \input{2dMultiplication.tex}
   \section{Plane rotations.}
      \input{2dRotations.tex}
   \section{Duality.}
     \input{dual.tex}
   \section{Vector product, dot product and wedge product.}
      \input{products.tex}
   \section{Reverse.}
      \input{reverse.tex}
   \section{Complex representations.}
      \input{complex.tex}
   %\section{Imaginary nature of the \R{3} pseudoscalar.}
      % depends on reverse
      %\input{R3PseudoscalarSquare.tex}
   \section{Multivector dot product.}
      \input{generalizedDot.tex}
      \subsection{Dot product of a vector and bivector}
         \input{generalizedDot_vectorBivector.tex}
      \subsection{Bivector dot product.}
         \input{generalizedDot_bivectorDot.tex}
      \subsection{Problems.}
         \input{distributionDotVectorBivector.tex}
   \section{Permutation within scalar selection.}
      \input{scalarSelectionPermutation.tex}
   \section{Multivector wedge product.}
      \input{generalizedWedge.tex}
      \subsection{Problems.}
         \input{generalizedWedgeProblems.tex}
         \input{r4nonzerobivectorwedgewithself.tex}
   \section{Projection and rejection.}
      \input{projectionAndRejection.tex}
   \section{Normal factorization of the wedge product.}
      \input{normalFactorization.tex}
   \section{The wedge product as an oriented area.}
      \input{wedgeProductArea.tex}
   \section{General rotation.}
      \input{generalRotation.tex}
   \section{Symmetric and antisymmetric vector sums.}
      \input{symmetricAndAntiSymmetricVectorSums.tex}
   \section{Reflection.}
      \input{reflection.tex}
   \section{Linear systems.}
      \input{solutionOfLinearSystem.tex}
  %\section{Orientation.}
  %   \input{working/orientation.tex}
   \section{A summary comparision.}
      \input{gacomparison.tex}
   \section{Problem solutions.}
      \shipoutAnswer
%}
