%
% Copyright © 2016 Peeter Joot.  All Rights Reserved.
% Licenced as described in the file LICENSE under the root directory of this GIT repository.
%
%----------------------------------------------------------------------------------------
\part{Geometric Algebra.}
   \chapter{Geometric Algebra.}
      \section{Prerequisites.}
         \input{prerequisites.tex}
         \subsection{Vector space.}
            \input{prereq_vectorspace.tex}
         \subsection{Basis, span and dimension.}
            \input{prereq_basisspandim.tex}
         \subsection{Standard basis, length and normality.}
            \input{prereq_standardbasis.tex}
      \section{Definitions}
         \subsection{Multivector space.}
            \input{multivector.tex}
         \subsection{Nomenclature.}
            \input{multivector_nomenclature.tex}
            \input{analysis_cavaet.tex}
      \section{Colinear vectors.}
         \input{colinearVectors.tex}
      \section{Normal vectors.}
         \input{normalVectors.tex}
      \section{2D multiplication table.}
         \input{2dMultiplication.tex}
      \section{Plane rotations.}
         \input{2dRotations.tex}
      \section{Duality.}
        \input{dual.tex}
      \section{Vector product, dot product and wedge product.}
         \input{products.tex}
      \section{Reverse.}
         \input{reverse.tex}
      \section{Complex representations.}
         \input{complex.tex}
      %\section{Imaginary nature of the \R{3} pseudoscalar.}
         % depends on reverse
         %\input{R3PseudoscalarSquare.tex}
      \section{Multivector dot product.}
         \input{generalizedDot.tex}
         \subsection{Dot product of a vector and bivector}
            \input{generalizedDot_vectorBivector.tex}
         \subsection{Bivector dot product.}
            \input{generalizedDot_bivectorDot.tex}
         \subsection{Problems.}
            \input{distributionDotVectorBivector.tex}
      \section{Permutation within scalar selection.}
         \input{scalarSelectionPermutation.tex}
      \section{Multivector wedge product.}
         \input{generalizedWedge.tex}
         \subsection{Problems.}
            \input{generalizedWedgeProblems.tex}
            \input{r4nonzerobivectorwedgewithself.tex}
      \section{Projection and rejection.}
         \input{projectionAndRejection.tex}
      \section{Normal factorization of the wedge product.}
         \input{normalFactorization.tex}
      \section{The wedge product as an oriented area.}
         \input{wedgeProductArea.tex}
      \section{General rotation.}
         \input{generalRotation.tex}
      \section{Symmetric and antisymmetric vector sums.}
         \input{symmetricAndAntiSymmetricVectorSums.tex}
      \section{Reflection.}
         \input{reflection.tex}
      \section{Linear systems.}
         \input{solutionOfLinearSystem.tex}
     %\section{Orientation.}
     %   \input{working/orientation.tex}
      \section{A summary comparision.}
         \input{gacomparison.tex}
      \section{Problem solutions.}
         \shipoutAnswer

   \chapter{Multivector calculus.}
      \section{Reciprocal frames.}
         \subsection{Motivation and definition.}
            \input{reciprocal.tex}
         \subsection{\R{2} reciprocal frame.}
            \input{reciprocal_R2.tex}
         \subsection{\R{3} reciprocal frame.}
            \input{reciprocal_R3.tex}
         \subsection{General reciprocal basis calculation.}
            \input{reciprocal_general.tex}
         \subsection{Problems.}
            \input{2dreciprocalMatrixCalculation.tex}
            \input{2subspaceR3reciprocalExample.tex}
      \section{Curvilinear coordinates.}
         \input{curvilinearDefined.tex}
         \input{curvilinearGradient.tex}

Many of the concepts are illuminated nicely by considering some examples.

         \subsection{Polar coordinates.}
            \input{2Dcylindrical.tex}

         \subsection{Spherical coordinates.}
            \input{curvilinearspherical.tex}

         \subsection{Toroidal coordinates.}
            \input{torusCenterOfMassParameterization.tex}

         \subsection{Problems.}
            \input{sphericaldot.tex}
            \input{volumeselection.tex}

%gabook: 31.1
%Also: Stokes chapter.
% Lots of examples there that should really be separated out from the stokes core content
%(now included here).
      \section{Green's theorem.}
         \input{greensTheorem.tex}
         %\subsection{Problems}
      \section{Stokes' theorem.}
         \subsection{Statement.}
            \input{statement.tex}
         \subsection{One parameter specialization of Stokes' theorem.}
            \input{oneparameter.tex}
         \subsection{Two parameter specialization of Stokes' theorem.}
            \input{twoparameter.tex}
         \subsection{Three parameter specialization of Stokes' theorem.}
            \input{threeparameter.tex}
         \subsection{Using scalar volume elements}
            \input{scalarVolumeElement.tex}
         \subsection{Problems}
            \input{stokesAndGreens.tex}
            \input{stokesTheoremCoreProblems.tex}

      \section{Fundamental theorem of geometric calculus.}
         %\subsection{Fundamental theorem of geometric calculus.}
         \input{fundamentalTheoremOfCalculus.tex}
      \section{Green's functions}
         \input{greensFunctionOverview.tex}
         \subsection{Green's function solutions.}
            \input{greensFunctionSolutions.tex}
         \subsection{Helmholtz equation.}
            \input{greensFunctionHelmholtz.tex}
% redundant:
%         \subsection{First order gradient.}
%            \input{greensFunctionGradient.tex}
         \subsection{First order Helmholtz equation.}
            \input{greensFunctionFirstOrderHelmholtz.tex}
         \subsection{Spacetime gradient.}
            \input{greensFunctionSpacetimeGradient.tex}
      \section{Helmholtz theorem.}
         \input{helmholtzDerviationMultivectorStatement.tex}
         \input{helmholtzDerviationMultivectorSolution.tex}
      \section{Problem solutions.}
         \shipoutAnswer

\part{Electromagnetism.}
   \chapter{Electromagnetism.}
      \section{Maxwell and Lorentz equations.}
         \input{freespace.tex}
         \subsection{Problems.}
            \input{landauEasy.tex}
      \section{Electrostatics.}
         \input{electrostatics.tex}
         \subsection{Enclosed charge.}
            \input{electrostatics_enclosedCharge.tex}
         \subsection{Electric potential.}
            \input{electrostatics_electricPotential.tex}
         \subsection{Inverting the gradient equations.}
            \input{electrostatics_invertingGradient.tex}
         \subsection{Poisson equation solution.}
            \input{electrostatics_poissonEquationSolution.tex}
         \subsection{Example: Straight line charge.}
            \input{linecharge.tex}
         \subsection{Example: Circular line charge.}
            \input{circularlinecharge.tex}
         \subsection{Problems.}
            \input{electrostaticsProblems.tex}
      \section{Magnetostatics.}
         \input{magnetostatics.tex}
         \subsection{Vector potential.}
            \input{magnetostatics_vectorPotential.tex}
         \subsection{Enclosed current density.}
            \input{magnetostatics_enclosedCurrentDensity.tex}
         \subsection{Enclosed current.}
            \input{magnetostatics_enclosedCurrent.tex}
         \subsection{Biot-Savart law.}
            \input{magnetostatics_biotSavart.tex}
         \subsection{Example.  Ampere's law for magnetic field between two current sources.}
            \input{amperesLawMagnetostaticsExample.tex}
      \section{Maxwell's equation (GA).}
         \input{isotropicMaxwells.tex}
      \subsection{Wave equation.}
         \input{waveequation.tex}
      \subsection{Continuity equations.}
         \input{continuity.tex}
      \section{Statics.}
         \subsection{Inverting the Maxwell statics equation.}
            \input{statics.tex}
         \subsection{Example.  Infinite line charge and current.}
            \input{statics_infiniteLineCharge.tex}
         \subsection{Example.  Infinite planar charge and current.}
            \input{statics_infinitePlane.tex}
         \subsection{Example.  Field of a ring of charge or current density.}
            \input{ringField.tex}
      \section{Dynamics.}
         \subsection{Inverting Maxwell's equation.}
            \input{jefimenkosEquations.tex}
      \section{Energy and Momentum.}
         \subsection{Field energy and momentum density and the energy-momentum tensor.}
            \input{poyntingF.tex}
         \subsection{Poynting's theorem.}
            \input{poyntingTheorem.tex}
         \subsection{Relation to Lorentz force.}
            \input{poyntingLorentzForce.tex}
         \subsection{Examples: Some static fields.}
            \input{poyntingStaticExamples.tex}
         \subsection{Complex energy and power.}
            \input{poyntingFComplexPower.tex}
      \section{Plane waves.}
         \input{planewaves.tex}
      \section{Polarization.}
         \subsection{Plane wave.}
            \input{polarization.tex}
         \subsection{Circular polarization basis.}
            \input{polarization_circular.tex}
         \subsection{Linear polarization.}
            \input{polarization_linearPolarization.tex}
         \subsection{Other phase dependence and energy momentum.}
            \input{polarization_phaseAndEnergyMomentum.tex}
         \subsection{Elliptical parameterization.}
            \input{polarization_elliptical.tex}
         \subsection{Pseudoscalar imaginary.}
            \input{polarization_pseudoscalarImaginary.tex}
         \subsection{Problems.}
            \input{polarization_problems.tex}
      \section{Transverse fields in a waveguide.}
         \input{transverseField.tex}
      \section{Boundary value conditions.}
         \input{boundarySurfaceSources.tex}
      \section{Multivector potential.}
         \subsection{General potential representation.}
            \input{potentialSection.tex}
         \subsection{Electric sources.}
            \input{potentialSection_electric.tex}
         \subsection{Magnetic sources.}
            \input{potentialSection_magnetic.tex}
         \subsection{Far field.}
            \input{potentialSection_farfield.tex}
         \subsection{Gauge transformations}
            \input{potentialSection_gauge.tex}
         \subsection{Lorenz gauge}
            \input{potentialSection_LorenzGauge.tex}
      \section{Lorentz force.}
         \subsection{GA statement.}
            \input{lorentzForce.tex}
         \subsection{Constant magnetic field.}
            \input{lorentzForce_constantMagnetic.tex}
      \section{Dielectric and magnetic media.}
         \input{inMatter.tex}
%   \section{Radiation and scattering}
%TODO.
   %   \subsection{Problem solutions}
   %      \shipoutAnswer

%\part{First draft fragments to rewrite.}
%   \chapter{Maxwell's equations}
%      \input{maxwellsEquations.tex}
%      \section{Potentials}
%         \input{potentialSection.tex}
%      \section{Problem solutions}
%         \shipoutAnswer
%%   \chapter{Time harmonic fields}
%%      \section{Frequency domain}
%%         \input{../frequencydomain/frequencydomainMaxwells.tex}
%%      \section{Plane waves}
%%         \input{frequencydomainPlaneWaves.tex}
%%      \section{Problem solutions}
%%         \shipoutAnswer
%   \chapter{Polarization}
%      \input{working/ellipticalWavesPhasor.tex}
%      \section{Problem solutions}
%         \shipoutAnswer
%%</TEMPORARY COMMENT OUT FOR QUICKER BUILD.>
