%
% Copyright © 2018 Peeter Joot.  All Rights Reserved.
% Licenced as described in the file LICENSE under the root directory of this GIT repository.
%
%{
Being able to compute the generalized dot product of two bivectors will also have a number of applications.
When those bivectors are wedge products, there is a useful distribution identity for this dot product.

\maketheorem{Dot product distribution over wedge products.}{thm:generalizedDot:bivectorDot}{
Given two sets of wedge products \( \Ba \wedge \Bb \), and \( \Bc \wedge \Bd \), their dot product is
\begin{equation*}
(\Ba \wedge \Bb) \cdot (\Bc \wedge \Bd)
=
\lr{ (\Ba \wedge \Bb) \cdot \Bc} \cdot \Bd
=
(\Bb \cdot \Bc) (\Ba \cdot \Bd)
-(\Ba \cdot \Bc)( \Bb \cdot \Bd).
\end{equation*}
} % theorem
\begin{proof}
To prove this, select the scalar grade of the product \( (\Ba \wedge \Bb) (\Bc \wedge \Bd) \)
\begin{equation}\label{eqn:generalizedDot_bivectorDot:1160}
\begin{aligned}
(\Ba \wedge \Bb) (\Bc \wedge \Bd)
&= (\Ba \wedge \Bb) (\Bc \Bd - \Bc \cdot \Bd) \\
&= (\Ba \wedge \Bb) \Bc \Bd - (\Ba \wedge \Bb) (\Bc \cdot \Bd).
\end{aligned}
\end{equation}

The second term, a bivector, is not of interest since it will be killed by the scalar selection operation.
The remainder can be expanded in grades, first making use of the fact that a bivector-vector product has only
grade 1 and 3 components
\begin{equation}\label{eqn:generalizedDot_bivectorDot:1180}
(\Ba \wedge \Bb) \Bc
=
(\Ba \wedge \Bb) \cdot \Bc
+ \gpgradethree{ (\Ba \wedge \Bb) \Bc }.
\end{equation}

Multiplication of the trivector term by \( \Bd \) produces a grade \((2,4)\)-multivector which can be ignored.
The product
of \( (\Ba \wedge \Bb) \cdot \Bc \), a vector, with \( \Bd \) is a grade \((0,2)\)-multivector, of which only the scalar grade is of interest.
That is
\begin{equation}\label{eqn:generalizedDot_bivectorDot:1220}
\begin{aligned}
(\Ba \wedge \Bb) \cdot (\Bc \wedge \Bd)
&= \gpgradezero{ (\Ba \wedge \Bb) (\Bc \wedge \Bd) } \\
&= ((\Ba \wedge \Bb) \cdot \Bc )\cdot \Bd.
\end{aligned}
\end{equation}

To complete the proof, we apply \cref{thm:generalizedDot:wedgeDotDistribution}
\begin{equation}\label{eqn:generalizedDot_bivectorDot:1240}
\begin{aligned}
((\Ba \wedge \Bb) \cdot \Bc )\cdot \Bd
&= \biglr{ \Ba (\Bb \cdot \Bc) - \Bb (\Ba \cdot \Bc) } \cdot \Bd \\
&= (\Ba \cdot \Bd) (\Bb \cdot \Bc) - (\Bb \cdot \Bd) (\Ba \cdot \Bc).
\end{aligned}
\end{equation}
\end{proof}

Identity 
\cref{eqn:generalizedDot_bivectorDot:1240}
has the following \R{3} cross product equivalent.
\maketheorem{Dot products of wedges as cross products.}{thm:generalizedDot:bivectorDotAsCrossProducts}{
The dot product of two \R{3} wedge products can be expressed as cross products
\begin{equation*}
(\Ba \wedge \Bb) \cdot (\Bc \wedge \Bd)
=
-(\Ba \cross \Bb) \cdot (\Bc \cross \Bd).
\end{equation*}
} % theorem
\begin{proof}
This follows by scalar grade selection
\begin{equation}\label{eqn:generalizedDot_bivectorDot:1140}
\begin{aligned}
(\Ba \wedge \Bb) \cdot (\Bc \wedge \Bd)
&= \gpgradezero{ (\Ba \wedge \Bb) (\Bc \wedge \Bd) } \\
&= \gpgradezero{ I (\Ba \cross \Bb) I (\Bc \cross \Bd) } \\
&= -(\Ba \cross \Bb) \cdot (\Bc \cross \Bd).
\end{aligned}
\end{equation}
\end{proof}

\begin{tablelabelbox}[tabularx={X||Y|Y}]{Comparison of distribution identities.}{label=tab:gen:distribution}
             & Geometric algebra & Vector algebra
\\ \hline
vector-bivector dot product (
triple cross product) & \( \Ba \cdot ( \Bb \wedge \Bc ) \) & \( (\Bb \cross \Bc) \cross \Ba \)
\\ \hline
bivector dot product (dot product of cross products) &
\( (\Ba \wedge \Bb) \cdot (\Bc \wedge \Bd) \)
&
\( -(\Ba \cross \Bb) \cdot (\Bc \cross \Bd) \)
\\ \hline
\end{tablelabelbox}

%}
