%
% Copyright © 2016 Peeter Joot.  All Rights Reserved.
% Licenced as described in the file LICENSE under the root directory of this GIT repository.
%
%{
The conventional proof of Helmholtz's theorem uses the Green's function for the (second order) Helmholtz operator.
Armed with a vector valued Green's function for the gradient, a first order proof is also possible.
As illustrations of the geometric integration theory developed in this chapter, both
strategies will be applied here to this problem.

In either case, we start by forming an even grade multivector (gradient) equation containing both the dot and cross product contributions
\begin{equation}\label{eqn:helmholtzDerviationMultivectorSolution:60}
\spacegrad \BM
= \spacegrad \cdot \BM + I \spacegrad \cross \BM
= s + I \BC.
\end{equation}

\paragraph{First order proof.}

For the first order case, we
perform a grade one selection of \cref{lemma:greensFunctionOverview:420}, setting
\( F = \BM \) where \( G \) is the Green's function for the gradient given by
\cref{eqn:greensFunctionFirstOrderHelmholtz:900}.  The proof follows directly

\begin{equation}\label{eqn:helmholtzDerviationMultivectorSolution:820}
\begin{aligned}
M(\Bx)
&= - \int_V \lr{ G(\Bx, \Bx') \lspacegrad' } \BM(\Bx') dV' \\
&= \int_V \gpgradeone{G(\Bx, \Bx') \lr{ \rspacegrad' \BM(\Bx') }} dV'
-
\int_{\partial V} \gpgradeone{ G(\Bx, \Bx') \ncap' \BM(\Bx') } dA' \\
&=
\int_V
\inv{4 \pi \Norm{\Bx - \Bx'}^3 }
\gpgradeone{ (\Bx - \Bx') \lr{ s(\Bx') + I \BC(\Bx') }} dV' \\
&\quad -
\int_{\partial V}
\inv{4 \pi \Norm{\Bx - \Bx'}^3 }
\gpgradeone{ (\Bx - \Bx') \ncap' \BM(\Bx') } dA' \\
&=
\int_V
\inv{4 \pi \Norm{\Bx - \Bx'}^3 }
\lr{ (\Bx - \Bx') s(\Bx') - (\Bx - \Bx') \cross \BC(\Bx') } dV' \\
&\quad -
\int_{\partial V}
\inv{4 \pi \Norm{\Bx - \Bx'}^3 }
\gpgradeone{ (\Bx - \Bx') \ncap' \BM(\Bx') } dA'.
\end{aligned}
\end{equation}
If \( \BM \) is well behaved enough that the boundary integral vanishes on an infinite surface, we see that \( \BM \) is completely specified by the divergence and the curl.
In general, the divergence and the curl, must also be supplemented by the value of vector valued function on the boundary.

Observe that the boundary integral has a particularly simple form for a spherical surface or radius \( R \) centered on \( \Bx' \).
Switching to spherical coordinates \( \Br = \Bx' - \Bx = R\, \rcap(\theta, \phi) \) where \( \rcap = (\Bx' - \Bx)/\Norm{\Bx' - \Bx} \) is the outwards normal, we have
\begin{equation}\label{eqn:helmholtzDerviationMultivectorSolution:840}
\begin{aligned}
-
\int_{\partial V} &
\inv{4 \pi \Norm{\Bx - \Bx'}^3 }
\gpgradeone{ (\Bx - \Bx') \ncap' \BM(\Bx') } dA' \\
&= \int_{\partial V} \frac{\BM(\Bx')}{4 \pi \Norm{\Bx - \Bx'}^2 } dA' \\
&= \inv{4\pi} \int_{\theta = 0}^\pi \int_{\phi = 0}^{2 \pi} \BM(R, \theta, \phi) \sin\theta d\theta d\phi.
\end{aligned}
\end{equation}
This is an average of \( \BM \) over the surface of the radius-\(R\) sphere surrounding the point \( \Bx \) where the field \( \BM \) is evaluated.

\paragraph{Second order proof.}

%Observe that the Laplacian of \( \BM \) is vector valued
%
%\begin{equation}\label{eqn:helmholtzDerviationMultivectorSolution:760}
%\spacegrad^2 \BM = \spacegrad s + I \spacegrad \BC.
%\end{equation}
%
%This means that \( \spacegrad \BC \) must be a bivector \( \spacegrad \BC = \spacegrad \wedge \BC \), or that \( \BC \) has zero divergence
%
%\begin{equation}\label{eqn:helmholtzDerviationMultivectorSolution:780}
%\spacegrad \cdot \BC = 0.
%\end{equation}

Again, we use \cref{eqn:helmholtzDerviationMultivectorSolution:60}
to discover the relation between the vector \( \BM \) and its divergence and curl.
\index{delta function}
The vector \( \BM \) can be expressed at the point of interest as a convolution with the delta function at all other points in space
\index{convolution}
\begin{equation}\label{eqn:helmholtzDerviationMultivectorSolution:80}
\BM(\Bx) = \int_V dV'\, \delta(\Bx - \Bx') \BM(\Bx').
\end{equation}

\index{Laplacian}
The Laplacian representation of the delta function in \R{3} is
\begin{equation}\label{eqn:helmholtzDerviationMultivectorSolution:100}
\delta(\Bx - \Bx') = -\inv{4\pi} \spacegrad^2 \inv{\Norm{\Bx - \Bx'}},
\end{equation}
so \( \BM \) can be represented as the following convolution
\begin{equation}\label{eqn:helmholtzDerviationMultivectorSolution:120}
\BM(\Bx) = -\inv{4\pi} \int_V dV'\, \spacegrad^2 \inv{\Norm{\Bx - \Bx'}} \BM(\Bx').
\end{equation}

%As noted in \cref{eqn:helmholtzDerviationMultivector:460} the Laplacian of a vector can be factored as
%
%\begin{equation}\label{eqn:helmholtzDerviationMultivectorSolution:140}
%\spacegrad^2 \Ba
%=
%\spacegrad (\spacegrad \cdot \Ba)
%-
%\spacegrad \cross (\spacegrad \cross \Ba).
%\end{equation}
%
%Note that the last term can be written in cross product notation using \( \Bc \cdot (\Ba \wedge \Bb) = -\Bc \cross (\Ba \cross \Bb) \) if desired.

Using this relation and proceeding with a few applications of the chain rule, plus the fact that \( \spacegrad 1/\Norm{\Bx - \Bx'} = -\spacegrad' 1/\Norm{\Bx - \Bx'} \), we find
%
%I previously posted a Geometric Algebra attack on the Helmholtz theorem.  Here is
%
%Here's a third way of deriving the Helmholtz theorem inversion relation.  This is a refinement of the traditional vector algebra solution that led to \cref{eqn:helmholtzDerviationMultivector:200}, that uses a factorization of the Laplacian directly, deferring any expansion in terms of dot and cross (or wedge) products until the very end.
%
%Starting from the first line of \cref{eqn:helmholtzDerviationMultivector:160}, we have
\begin{equation}\label{eqn:helmholtzDerviationMultivectorSolution:720}
\begin{aligned}
-4 &\pi \BM(\Bx) \\
&= \int_V dV'\, \spacegrad^2 \inv{\Norm{\Bx - \Bx'}} \BM(\Bx') \\
&= \gpgradeone{\int_V dV'\, \spacegrad^2 \inv{\Norm{\Bx - \Bx'}} \BM(\Bx')} \\
&= -\gpgradeone{\int_V dV'\, \spacegrad \lr{ \spacegrad' \inv{\Norm{\Bx - \Bx'}}} \BM(\Bx')} \\
&= -\gpgradeone{\spacegrad \int_V dV' \lr{
   \spacegrad' \frac{\BM(\Bx')}{\Norm{\Bx - \Bx'}}
   -\frac{\spacegrad' \BM(\Bx')}{\Norm{\Bx - \Bx'}}
   } } \\
&= -\gpgradeone{\spacegrad \int_{\partial V} dA'\,
   \ncap \frac{\BM(\Bx')}{\Norm{\Bx - \Bx'}}
    }
   +\gpgradeone{\spacegrad \int_V dV'
   \frac{s(\Bx') + I\BC(\Bx')}{\Norm{\Bx - \Bx'}}
    } \\
&= -\gpgradeone{\spacegrad \int_{\partial V} dA'\,
   \ncap \frac{\BM(\Bx')}{\Norm{\Bx - \Bx'}}
    }
   +\spacegrad \int_V dV'\,
   \frac{s(\Bx')}{\Norm{\Bx - \Bx'}}
   +\spacegrad \cdot \int_V dV'
   \frac{I\BC(\Bx')}{\Norm{\Bx - \Bx'}}.
\end{aligned}
\end{equation}

By inserting a no-op grade selection operation in the second step, the trivector terms that would show up in subsequent steps are automatically filtered out.
%the troublesome trivector term that shows up in my first purely Geometric Algebra
%attempt is eliminated.
This leaves us with a boundary term dependent on the surface and the normal and tangential components of \( \BM \).
Added to that is a pair of volume integrals that provide the unique dependence of \( \BM \) on its divergence and curl.
When the surface is taken to infinity, which requires \( \Norm{\BM}/\Norm{\Bx - \Bx'} \rightarrow 0 \), then the dependence of \( \BM \) on its divergence and curl is unique.

In order to express final result in traditional vector algebra form, a couple transformations are required.
The first is that
\begin{equation}\label{eqn:helmholtzDerviationMultivectorSolution:800}
\gpgradeone{ \Ba I \Bb } = I^2 \Ba \cross \Bb = -\Ba \cross \Bb.
\end{equation}

For the grade selection in the boundary integral, note that
\begin{equation}\label{eqn:helmholtzDerviationMultivectorSolution:740}
\begin{aligned}
\gpgradeone{ \spacegrad \ncap \BX }
&= \gpgradeone{ \spacegrad (\ncap \cdot \BX) } + \gpgradeone{ \spacegrad (\ncap \wedge \BX) } \\
&= \spacegrad (\ncap \cdot \BX) + \gpgradeone{ \spacegrad I (\ncap \cross \BX) } \\
&= \spacegrad (\ncap \cdot \BX) - \spacegrad \cross (\ncap \cross \BX).
\end{aligned}
\end{equation}

These give
%\begin{equation}\label{eqn:helmholtzDerviationMultivectorSolution:721}
\boxedEquation{eqn:helmholtzDerviationMultivectorSolution:721}{
\begin{aligned}
\BM(\Bx)
&=
\spacegrad \inv{4\pi} \int_{\partial V} dA'\, \ncap \cdot \frac{\BM(\Bx')}{\Norm{\Bx - \Bx'}}
-
\spacegrad \cross \inv{4\pi} \int_{\partial V} dA'\, \ncap \cross \frac{\BM(\Bx')}{\Norm{\Bx - \Bx'}} \\
&-\spacegrad \inv{4\pi} \int_V dV'
\frac{s(\Bx')}{\Norm{\Bx - \Bx'}}
+\spacegrad \cross \inv{4\pi} \int_V dV'
\frac{\BC(\Bx')}{\Norm{\Bx - \Bx'}}.
\end{aligned}
}
%\end{equation}
%}
