%
% Copyright © 2017 Peeter Joot.  All Rights Reserved.
% Licenced as described in the file LICENSE under the root directory of this GIT repository.
%
\index{reverse}
\input{reverseDefined.tex}

A useful relation for k-vectors that are composed entirely of products of orthogonal vectors exists.  We call such k-vectors blades

\index{blade}
\index{grade}
\index{\(\bigwedge^k\)}
\makedefinition{Blade.}{def:multiplication:blade}{
A product of \( k \) orthogonal vectors is called a k-blade, or a blade of grade \( k \).
A grade zero blade is a scalar.

The notation \( F \in \bigwedge^k \) is used in the literature to indicate that \( F \) is a blade of grade \( k \).
}

Any k-blade is also a k-vector, but not all k-vectors are k-blades.  For example in \R{4} the bivector
\(
\Be_{12} + \Be_{34} \)
is not a 2-blade, since it cannot be factored into orthogonal products, whereas any \R{3} bivector, such as
\( \Be_{12} + \Be_{23} + \Be_{31} \) is a blade (\cref{problem:normalFactorization:1}).
This will be relevant when formulating rotations since bivectors that are blades can be used to simply describe rotations or Lorentz boosts
\footnote{A rotation in spacetime.} whereas it is not easily possible to compute an exponential of a non-blade bivector argument.

\maketheorem{Reverse of k-blade.}{thm:reverse:kBlade}{
The reverse of a k-blade \( A_k = \Ba_1 \Ba_2 \cdots \Ba_k \) is given by
\begin{equation*}
A_k^\dagger = (-1)^{k(k-1)/2} A_k.
\end{equation*}
} % theorem

This can be proven by successive interchange of the factors
\begin{dmath}\label{eqn:reverse:81}
\begin{aligned}
A_k^\dagger
&= \Ba_k \Ba_{k-1} \cdots \Ba_1 \\
&= (-1)^{k-1} \Ba_1 \Ba_k \Ba_{k-1} \cdots \Ba_2 \\
&= (-1)^{k-1} (-1)^{k-2} \Ba_1 \Ba_2 \Ba_k \Ba_{k-1} \cdots \Ba_3 \\
&\qquad \vdots \\
&= (-1)^{k-1} (-1)^{k-2} \cdots (-1)^1 \Ba_1 \Ba_2 \cdots \Ba_k \\
&= (-1)^{k(k-1)/2} \Ba_1 \Ba_2 \cdots \Ba_k \\
&= (-1)^{k(k-1)/2} A_k,
% \qquad \qedmarker
\end{aligned}
\end{dmath}
which completes the proof.

A special, but important case, is the reverse of the \R{3} pseudoscalar, which is negated by reversion
\boxedEquation{eqn:reverse:103}{
I^\dagger = -I.
}

