%\section{Two dimensions}

\paragraph{Multiplication table}

In a 2D space most of the interesting vector products involve the unit bivector \( \Be_1 \Be_2 \).  A product of a spanning set of normal vectors for a space (or subspace) is called a pseudoscalar for that space (or subspace).  The particular pseudoscalar \( \Be_1 \Be_2 \) will be labelled \textbf{the pseudoscalar}.  Computation shows that multiplication with the pseudoscalar anticommutes with the 2D basis vectors

\begin{dmath}\label{eqn:SimpleProducts2:180}
\begin{aligned}
   \Be_1 \lr{ \Be_1 \Be_2 } &= \lr{ \Be_1 \Be_1 } \Be_2 = \Be_2 \\
   \lr{ \Be_1 \Be_2 } \Be_1 &= -\lr{ \Be_2 \Be_1 } \Be_1 = -\Be_2 \\
\end{aligned}
\end{dmath}

\begin{dmath}\label{eqn:SimpleProducts2:200}
\begin{aligned}
   \Be_2 \lr{ \Be_1 \Be_2 } &= \lr{ -\Be_1 \Be_2 } \Be_2 = -\Be_1 \\
   \lr{ \Be_1 \Be_2 } \Be_2 &= \Be_1 \lr{ \Be_2 \Be_2 } l= \Be_1 \\
\end{aligned}
\end{dmath}

Observe that in \R{2} the product of any basis vector with a pseudoscalar is normal to the original vector, which is also generally true for any vector in a 2D space.  Such a multiplication induces a \( \pi/2 \) rotation, the direction of which depends on the orientation of pseudoscalar, and upon whether the multiplication is performed from the left or the right.  A hint of the rotational nature of such a product can be gleamed by computing the square of the 2D pseudoscalar

\begin{dmath}\label{eqn:SimpleProducts2:220}
   \lr{ \Be_1 \Be_2 }^2
   =
   \Be_1 \Be_2
   \Be_1 \Be_2
   =
   \Be_1 \lr{ \Be_2
   \Be_1 } \Be_2
   =
   \Be_1 \lr{ -\Be_1
   \Be_2 } \Be_2
   =
   -\lr{ \Be_1 \Be_1 }
   \lr{ \Be_2 \Be_2 }
   = -1.
\end{dmath}

This unit bivector is seen to square to minus one like the imaginary in complex algebra.  The reader can confirm easily that this is generally true for any unit bivector \( \Be_i \Be_j, \, i \ne j \).
This is a very convienient fact, and allows ad-hoc construction of complex number like coordinate systems in any given planar subspace.

The products above are summarized in \cref{tab:SimpleProducts2:10}.
FIXME: make prettier and center.
% https://tex.stackexchange.com/a/135421/15
% https://tex.stackexchange.com/a/298109/15
% https://tex.stackexchange.com/a/112359/15

\captionedTable{2D Multiplication table.}{tab:SimpleProducts2:10}{
\begin{tabular}{|l||l|l|l|l|}
\hline
&
\( 1 \) & \( \Be_1 \) & \( \Be_2 \) & \( \Be_1 \Be_2 \) \\
\hline
\( 1 \) & \( 1 \) & \( \Be_1 \) & \( \Be_2 \) & \( \Be_1 \Be_2 \) \\
\hline
\( \Be_1\) & \( \Be_1 \) & \( 1 \) & \( \Be_1 \Be_2 \) & \( \Be_2 \)\\
\hline
\( \Be_2\) & \( \Be_2 \) & \( -\Be_1 \Be_2 \) & \( 1 \) & \( -\Be_1 \)\\
\hline
\( \Be_1 \Be_2\) & \( \Be_1 \Be_2 \) & \( -\Be_2 \) & \( \Be_1 \) & \( -1 \) \\
\hline
\end{tabular}
}

\paragraph{Computing the normal 2D}

Given a coordinate representation of an arbitrary vector in a 2D space

\begin{dmath}\label{eqn:SimpleProducts2:240}
   \Bx = \rho
\begin{bmatrix}
   \cos\theta \\
   \sin\theta \\
\end{bmatrix},
\end{dmath}

after counterclockwise rotation by \( \pi/2 \), the rotated coordinates are

\begin{dmath}\label{eqn:SimpleProducts2:260}
\Bx'
=
\begin{bmatrix}
   0 & -1 \\
   1 & 0 \\
\end{bmatrix}
\begin{bmatrix}
   \cos\theta \\
   \sin\theta \\
\end{bmatrix}
=
\rho
\begin{bmatrix}
   -\sin\theta \\
   \cos\theta
\end{bmatrix}.
\end{dmath}

Expressing this vector in terms of the standard basis

\begin{dmath}\label{eqn:SimpleProducts2:280}
   \Bx = \rho \lr{ \Be_1 \cos\theta + \Be_2 \sin\theta },
\end{dmath}

the same rotation can be observed by right multiplication by the pseudoscalar.

\begin{dmath}\label{eqn:SimpleProducts2:300}
\Bx'
= \rho \lr{ \Be_1 \cos\theta + \Be_2 \sin\theta } \Be_1 \Be_2
= \rho \lr{ \Be_2 \cos\theta - \Be_1 \sin\theta }.
\end{dmath}

FIXME: illustration.

A similar left pseudoscalar multiplication induces a clockwise rotation.
(\cref{problem:left2dimaginarymultiplication:1}).

Just as the imaginary rotates complex numbers, the 2D pseudoscalar rotates vectors, with the cavaet that one must be careful about the order that this multiplication is performed.

\paragraph{Complex numbers and rotations}

The 2D pseudoscalar has been seen to have the characteristics of the complex imaginary.  This analogy can be extended by noting that the following multivector

\begin{dmath}\label{eqn:SimpleProducts2:320}
z = x + \Be_1 \Be_2 y,
\end{dmath}

provides an isomorphic representation of a complex number.  This can be made obvious by introduces the lable \( i = \Be_1 \Be_2 \) for the pseudoscalar, since it has the desired characteristics of the imaginary.  

Of particular interest is the complex exponential \( e^{i \theta} \).  Because \( i \) commutes with itself and any scalars Euler's formula

\begin{dmath}\label{eqn:SimpleProducts2:340}
e^{i \theta} = \cos\theta + i \sin\theta,
\end{dmath}

applies equally well with this GA representation of complex numbers.  Having seen that right multiplication of vectors with with the pseudoscalar \( i \) had the effect of inducing a \( \pi/2 \) counterclockwise rotation, it is reasonable to expect that this form of complex exponential will generally rotate a 2D vector.  With

\begin{dmath}\label{eqn:SimpleProducts2:360}
\Bx
= x \Be_1 + y \Be_2,
\end{dmath}

multiplication with \( e^{i\theta} \) gives

\begin{dmath}\label{eqn:SimpleProducts2:380}
\Bx'
= \lr{ x \Be_1 + y \Be_2 } e^{i\theta}
= \lr{ x \Be_1 + y \Be_2 } \lr{ \cos\theta + i \sin\theta }
= \lr{ x \Be_1 + y \Be_2 } \lr{ \cos\theta + \Be_1 \Be_2 \sin\theta }
=
x \cos\theta \Be_1
+
x \sin\theta \Be_2
+
y \cos\theta \Be_2
-
y \sin\theta \Be_1
=
\lr{ x \cos\theta - y \sin\theta } \Be_1
+
\lr{ x \sin\theta + y \sin\theta } \Be_2.
\end{dmath}

Observe that this matches the rotation by \( \theta \) of the coordinates

\begin{dmath}\label{eqn:SimpleProducts2:400}
\begin{bmatrix}
   \cos\theta & - \sin\theta \\
   \sin\theta &   \cos\theta \\
\end{bmatrix}
\begin{bmatrix}
   x \\
   y
\end{bmatrix}
=
\begin{bmatrix}
   x \cos\theta  - y \sin\theta \\
   x \sin\theta  + y \cos\theta \\
\end{bmatrix},
\end{dmath}

with both transformations inducing a transformation

\begin{dmath}\label{eqn:SimpleProducts2:420}
\begin{aligned}
x &\rightarrow x \cos\theta  - y \sin\theta \\
y &\rightarrow x \sin\theta  + y \cos\theta \\
\end{aligned}.
\end{dmath}

This
rotation can also be encoded using a right multiplication with the ``conjugate'' complex exponential (\cref{problem:sandwichMult:1}).

\begin{dmath}\label{eqn:SimpleProducts2:440}
\Bx'
= e^{-i\theta} \lr{ x \Be_1 + y \Be_2 }.
\end{dmath}

It is also possible to encode this rotation in a half angle ``sandwich'' form that may be familiar to students of quaterions or of quantum mechanics

\begin{dmath}\label{eqn:SimpleProducts2:460}
\Bx'
= e^{-i\theta/2} \lr{ x \Be_1 + y \Be_2 } e^{i \theta/2}.
\end{dmath}

This follows from the fact that this complex exponential exhibits conjugate commutation behaviour (\cref{problem:sandwichMult:1})

\begin{dmath}\label{eqn:SimpleProducts2:480}
\begin{aligned}
   \Be_1 e^{i\theta} &= e^{-i\theta} \Be_1 \\
   \Be_2 e^{i\theta} &= e^{-i\theta} \Be_2 \\
\end{aligned}.
\end{dmath}

Later, we will see that this latter half angle rotation operator will be the desired form of a general rotation in \R{N}.  This generalized rotational structure that follows that such a complex exponential commutes with any vectors that lie outside of the plane (\cref{problem:normalMult:1}).

\paragraph{Vector product.}

The product of two 2D vectors is illustrative.  Let

\begin{dmath}\label{eqn:SimpleProducts2:500}
\begin{aligned}
   \Bx &= x_1 \Be_1 + x_2 \Be_2 \\
   \By &= y_1 \Be_1 + y_2 \Be_2
\end{aligned},
\end{dmath}

and consider their product
\begin{dmath}\label{eqn:SimpleProducts2:520}
\Bx \By
=
\lr{ x_1 \Be_1 + x_2 \Be_2 }
\lr{ y_1 \Be_1 + y_2 \Be_2 }
=
x_1 y_1 \Be_1 \Be_1 + x_2 y_1 \Be_2 \Be_1
+
x_1 y_2 \Be_1 \Be_2 + x_2 y_2 \Be_2 \Be_2
=
x_1 y_1
- x_2 y_1 \Be_1 \Be_2
+ x_1 y_2 \Be_1 \Be_2
+ x_2 y_2
=
x_1 y_1 + x_2 y_2
+ \lr{ x_1 y_2 - x_2 y_1 } \Be_1 \Be_2.
\end{dmath}

The vector product has a scalar and a bivector component, where the scalar component is completely symmetric, whereas the bivector component is completely antisymmetric.  Observe that the scalar selection from the vector product is the dot product.  A name is required for the bivector (grade-2) selection from this product, and it is called the wedge product \( \Bx \wedge \By \), an operation that has to be examined in more depth.  With such a definition, the vector product can be written as

%\begin{dmath}\label{eqn:SimpleProducts2:540}
\boxedEquation{eqn:SimpleProducts2:540}{
\Bx \By
= \Bx \cdot \By + \Bx \wedge \By,
}
%\end{dmath}

where

\begin{dmath}\label{eqn:SimpleProducts2:560}
\begin{aligned}
\Bx \cdot \By &\equiv \gpgradezero{\Bx\By} = x_1 y_1 + x_2 y_2 \\
\Bx \wedge \By &\equiv \gpgradetwo{\Bx\By} =
\begin{vmatrix}
   x_1 & x_2 \\
   y_1 & y_2
\end{vmatrix}
   \Be_1 \Be_2. \\
\end{aligned}
\end{dmath}

Because the wedge product is completely antisymmetric, it must be true that

\begin{dmath}\label{eqn:SimpleProducts2:580}
\By \wedge \Bx = -\Bx \wedge \By,
\end{dmath}

so
\begin{dmath}\label{eqn:SimpleProducts2:600}
\By \Bx
= \By \cdot \Bx + \By \wedge \Bx
= \Bx \cdot \By - \Bx \wedge \By.
\end{dmath}

Taken together \cref{eqn:SimpleProducts2:540} and \cref{eqn:SimpleProducts2:600} allow for a construction of a coordinate free form of both the dot and wedge products

\begin{dmath}\label{eqn:SimpleProducts2:620}
\begin{aligned}
\Bx \cdot \By   &= \inv{2}\lr{ \Bx \By + \By \Bx } \\
\Bx \wedge \By  &= \inv{2}\lr{ \Bx \By - \By \Bx }.
\end{aligned}
\end{dmath}

These highlight the symmetric and antisymmetric nature of the respective dot and wedge products.  Some authors will use \cref{eqn:SimpleProducts2:620} as the definitions of the dot and wedge products instead of defining them in terms of grade selection.  Grade selection is preferred here since it allows for a generalization of the wedge product to multiple vectors in higher degree spaces in a particularily simple way, and also allows for the generalization of the dot and wedge products with higher order geometric structures to be discussed.

\paragraph{Area}

It was previously claimed that the pseudoscalar \( \Be_1 \Be_2 \) could be interpretted as an oriented (signed) area.  Because \( \Be_1 \cdot \Be_2 = 0 \), this vector product is also equal to the wedge

\begin{dmath}\label{eqn:SimpleProducts2:640}
\Be_1 \Be_2 = \Be_1 \cdot \Be_2 +
\Be_1 \wedge \Be_2
=
\Be_1 \wedge \Be_2.
\end{dmath}

Recall that the area of the parallopiped spanned by two vectors in a two dimensional space is given by the absolute value of

\begin{dmath}\label{eqn:SimpleProducts2:660}
\begin{vmatrix}
   x_1 & x_2 \\
   y_1 & y_2
\end{vmatrix},
\end{dmath}

a factor that was also found in the coordinate expansion of the wedge product (\cref{eqn:SimpleProducts2:560}).  It is therefore natural to interpret the wedge product as an oriented area.
Without reference to a generating pair of vectors, a general bivector will be represented as an oriented area, where the orientation is depicted using a loop, as in
FIXME:figure.
In higher dimensonal spaces, the bivector factor not only encodes a sign for this area, but also its orientation in space.  The wedge product will be seen to encode that orientation without introducing a normal direction for the spanning plane, a nice feature in higher dimensional spaces where a single normal direction is ambiguous.

\paragraph{Projection and rejection}

FIXME: cartesian
FIXME: polar
